---
title: "Reproducible Research: Week 2 Project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Read in the data and clean

```{r}
file <- "activity.csv"
activityraw <- read.csv(file)
activityraw$date<- as.Date(activityraw$date)
```

## What is the mean total number of steps taken per day?
Calculate the total number of steps taken per day

Plot on a histogram

```{r activityraw, echo=TRUE}
library(dplyr)
stepsbydate <- activityraw %>% 
  group_by(date) %>%
  summarize(datesteps = sum(steps)) %>%
  na.omit()
hist(stepsbydate$datesteps, xlab = "Steps Each Day", main = "Histogram of Steps Taken Each Day", breaks = 20)
```

The mean and median of the total number of steps taken per day:

```{r activityraw2, echo=TRUE}
mean(stepsbydate$datesteps)
median(stepsbydate$datesteps)
```

## What is the average daily activity pattern?

```{r activitypattern, echo=TRUE}
activitypattern <- activityraw %>%
  na.omit() %>%
  group_by(interval) %>%
  summarize(avgsteps = mean(steps))
library(ggplot2)
ggplot(activitypattern, aes(x = interval, y = avgsteps)) +
  geom_line()

```

## Which 5-minute interval contains the maximum number of steps?

```{r intervals, echo=TRUE}
activitypattern[which(activitypattern$avgsteps == max(activitypattern$avgsteps)),]
```

## Imputing missing values

How many observations in the dataset have missing values?

```{r missingvalues, echo = TRUE}
sum(is.na(activityraw))
```

Fill in missing values. 

_I have chosen to use the mean for that five minute interval_

```{r replacemissing, echo = TRUE}
replacewithmean <- function(x) replace(x, is.na(x), mean(x, na.rm=TRUE))
intervalmean <- activityraw %>%
  group_by(interval) %>%
  mutate(steps = replacewithmean(steps))
```

## Make a histogram of the new data.

```{r newhistogram, echo=TRUE}
intervalstepsbydate <- intervalmean %>% 
  group_by(date) %>%
  summarize(datesteps = sum(steps))
hist(intervalstepsbydate$datesteps, xlab = "Steps Each Day", main = "Histogram of Steps Taken Each Day", breaks = 20)
```

Calculate the mean and median total number of steps taken per day

```{r newmeanmedian, echo = TRUE}
mean(intervalstepsbydate$datesteps)
median(intervalstepsbydate$datesteps)
```

## Are there differences in activity patterns between weekdays and weekends?

Create a factor variable for "weekday" and "weekend"

_I chose to use the dataset with imputed step counts_

```{r patterns, echo = TRUE} 
weekdays <- intervalmean
weekdays$date <- weekdays(weekdays$date)
weekdays$date <- ifelse(weekdays$date == "Saturday" | weekdays$date == "Sunday", "weekend","weekday")
stepsbyday <- weekdays %>%
  group_by(date, interval) %>%
  summarize(meansteps = mean(steps))
ggplot(stepsbyday, aes(x = interval, y = meansteps)) +
  geom_line() +
  xlab("Interval") +
  ylab("Mean of Steps") +
  facet_grid(~date)
```

